MADA Data Analysis Project

### 2021-11-24 products/manuscript/Manuscript.Rmd
Pre holiday Push
Push current updates to see what else can be done  for manuscript

Applying MADA learned techniques to a RT-PCR dataset to determine the effectiveness of sample enhancement strategies for environmental surface sampling of low bio-burden viruses.

@@

##  Background
The global Coronavirus Disease 2019 (COVID-19) pandemic continues to persist due to the timing of an available vaccine, varying adherence to COVID-19 public health safety recommendations, and the emergence of SARS-CoV-2 (SC2) variants. Environmental reservoirs are a reported, indirect transmission route for viral pathogens [@Marcenac2021-sz]. Although many viruses have a low infectious dose, they can be difficult to detect due to the low bioburden in the environment. The ability to detect viral pathogens is critical to identifying hidden reservoirs which can inform infection control [@Paton2021-jc]. Culture-based methods can be labor and time intensive. The use and limitations of commercial “target capture” (TC) technologies are varied compared to molecular approaches utilizing automated extraction. The objective of this study is improve outbreak response through rapid pathogen detection by evaluating a TC technology’s ability to capture SC2 from complex matrices.

@@

## Data aquisition
Reports per each rRT-PCR run were generated and quality checked on their respective machines before transfer to a Teams repository. Summary data was selected from the reports and added to additional environmental data gathered on based on the extraction date of the sample.

@@

Atempt to do as [@Brandolini2021-ja]

@@

Summary table 
scatter plot of SDs based on dilution factors
```{r}
knitr::include_graphics("../../results/ALL-SD-Plot.png")
```
#### Phase 1

@@

```{r}
knitr::include_graphics("../../results/P2-SD-Plot.png")
```
#### Phase 2 Comparing A and B Tool Recovery 
```{r}
knitr::include_graphics("../../results/ASK-P2P1-Tools.png")
```
Tool A has an average Ct recovery higher than B. This makes sense as the surfaces A recovered from is 2 logs higher (-1) than that of B (-3). A log drop is ~3Ct values we can see from the graphic that it trends as predicted. Suggests that

#### Phase 2 Comparing Method Serial Dilutions (CDC vs NTP) with standard input of 300uL
```{r}
knitr::include_graphics("../../results/ASK-P2P2-Methods.png")
```
Serial dilution data between CDC and NTP. Noting that all SD are in the confirmed realm of under 27 Ct Value CDC method is better than NTP at -1 concentration of the SD. the -2 shows little difference between method recovery Ct values. NTP appears to remain better than CDC at -3 and -4 levels. Accounting for the SE looks like both methods are the same until the -2 dilution. Might be significant difference at the -4 dilution.

#### Phase 2 Comparing Methods with standard input of 300uL
```{r}
knitr::include_graphics("../../results/ASK-P2P3-Methods0300.png")
```

#### Phase 2 Method extraction comparisons with all inputs and tools
```{r}
knitr::include_graphics("../../results/ASK-P2P4-MethodsALL.png")
```

#### Phase 2 Nanoparticles only - Compare detectors with dilution and input in Serial Dilutions
```{r}
knitr::include_graphics("../../results/ASK-P2P5-NTP-SD.png")
```

#### Phase 2 Nanoparticles only - All tools at 300 and 1000ul input levels
```{r}
knitr::include_graphics("../../results/ASK-P2P6-NTP.png")
```

@@

## Summary and Interpretation
Took

## Strengths and Limitations
Need more replicates, but this is a time consuming processes just to get the data.

>>>>

Applying MADA learned techniques to a RT-PCR data set to determine the effectiveness of sample enhancement strategies for environmental surface sampling of low bio-burden viruses.

@@

## Background

The global Coronavirus Disease 2019 (COVID-19) pandemic continues to persist due to the timing of an available vaccine, varying adherence to COVID-19 public health safety recommendations, and the emergence of SARS-CoV-2 (SC2) variants. Environmental reservoirs are a reported, indirect transmission route for viral pathogens [@Marcenac2021-sz]. Although many viruses have a low infectious dose, they can be difficult to detect due to the low bio-burden in the environment. The ability to detect viral pathogens is critical to identifying hidden reservoirs which can inform infection control [@Paton2021-jc]. Culture-based methods can be labor and time intensive. The use and limitations of commercial "target capture" (TC) technologies are varied compared to molecular approaches utilizing automated extraction. The objective of this study is improve outbreak response through rapid pathogen detection by evaluating a TC technology's ability to capture SC2 from complex matrices.

@@

## Data aquisition

Reports per each rRT-PCR run were generated and quality checked on their respective machines before transfer to a Teams repository. Summary data was selected from the reports and added to additional environmental data gathered on based on the extraction date of the sample.

@@

@@

Summary table scatter plot of SDs based on dilution factors

```{r}
knitr::include_graphics("../../results/ALL-SD-Plot.png")
```

#### Phase 1

@@

```{r}
knitr::include_graphics("../../results/P2-SD-Plot.png")
```

#### Phase 2 Comparing A and B Tool Recovery

```{r}
knitr::include_graphics("../../results/ASK-P2P1-Tools.png")
```

Tool A has an average Ct recovery higher than B. This makes sense as the surfaces A recovered from is 2 logs higher (-1) than that of B (-3). A log drop is \~3Ct values we can see from the graphic that it trends as predicted. Suggests that

#### Phase 2 Comparing Method Serial Dilutions (CDC vs NTP) with standard input of 300uL

```{r}
knitr::include_graphics("../../results/ASK-P2P2-Methods.png")
```

Serial dilution data between CDC and NTP. Noting that all SD are in the confirmed realm of under 27 Ct Value CDC method is better than NTP at -1 concentration of the SD. the -2 shows little difference between method recovery Ct values. NTP appears to remain better than CDC at -3 and -4 levels. Accounting for the SE looks like both methods are the same until the -2 dilution. Might be significant difference at the -4 dilution.

#### Phase 2 Comparing Methods with standard input of 300uL

```{r}
knitr::include_graphics("../../results/ASK-P2P3-Methods0300.png")
```

#### Phase 2 Method extraction comparisons with all inputs and tools

```{r}
knitr::include_graphics("../../results/ASK-P2P4-MethodsALL.png")
```

#### Phase 2 Nanoparticles only - Compare detectors with dilution and input in Serial Dilutions

```{r}
knitr::include_graphics("../../results/ASK-P2P5-NTP-SD.png")
```

#### Phase 2 Nanoparticles only - All tools at 300 and 1000ul input levels

```{r}
knitr::include_graphics("../../results/ASK-P2P6-NTP.png")
```

@@

## Summary and Interpretation

Took

## Strengths and Limitations

Need more replicates, but this is a time consuming processes just to get the data.
